Universals of word order result from optimization of grammars for efficient communication

### 2019-04-07 second-draft/3-paper-names.tex
revising

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic

@@

Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study~\cite{lehmann1973structural, vennemann1974theoretical,jackendoff1977x,frazier1985syntactic,chomsky1988language, dryer1992greenbergian, hawkins1994performance}.
A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.

@@

Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between the Forces of \emph{Unification} and \emph{Diversification}.
The Force of Unification is a pressure to reduce the complexity of the language to make production and processing as easy as possible.
Conversely, the Force of Diversification favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
%Minimizing speaker and listener effort represent two opposing forces:
%A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
%A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
These two forces act in opposing directions:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required in order to provide higher informativity.
The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward, schwartz1997dispersion}.

@@

While it has been suggested that these ideas should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult in the domain of grammar, as this requires both large amounts of data representative of language use across languages and cultures, and computational methods for estimating the efficiency of entire languages.
%TODO somewhere start calling this Efficiency
%TODO explain why surprisal. some production/speaker work. (plan reuse)
%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
%This theory blabla Horn RSA
%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
In this work, we address these challenges by combining large-scale text data from 51 languages with state-of-the-art machine learning techiques to estimate the communicative efficiency of grammar.
We apply this approach to the problem of explaining Greenberg's word order correlation universals.
We first show that the word order of languages has evolved to optimize efficiency.
We then show that efficiency optimization accounts for the Greenbergian word order correlations, and makes fine-grained per-language predictions about individual syntactic patterns.
%We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.

@@

\begin{table}
	\begin{center}
%		\setlength\tabcolsep{1.5pt}
\begin{tabular}{|c|ll|ll|}
	\hline
	%&
%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
%	&	Verb & Object     & Object & Verb    \\ \hline\hline
	& kataba & rasaa'il 	&tegami-o & kaita \\
	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    sadiiq       &	tomodachi & ni \\
	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&kaana    &    sadiiq         &	tomodachi & datta \\
	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak & udesho \\
	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& sadiiq  &    John    & 	John no & tomodachi \\
	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uha       & 	anata-ga yonda & hon \\
	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tusil        & 	toochaku suru & koto \\
	&	\emph{that} & \emph{she arrives}  &	\emph{has.arrived} & \emph{that} \\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'ila lmadrasa     & 	gakkoo ni & ittekimashita \\
	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'uriid    &    'an 'ughaadir         & 	yom & itai \\
	& \emph{wants}   &  \emph{to leave}  &	\emph{to leave} & \emph{want} \\ \hline \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples. Based on Dryer \cite{dryer1992greenbergian}, see Materials and Methods.}\label{table:corr-dryer}
\end{table}

@@

%In Information Theory, the usefulness of a communication channel 
The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
%This idea formalizes Zipf's forces of Unification and Diversification
Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:
%The \key{efficiency} of a language is a weighted combination
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Informativity}} - \lambda R_\textit{Cost}
\end{equation}
with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).

@@

%In our model, both aspects of efficiency will depend on the grammar:
Grammars differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
This reflects the \key{parseability} of the language.
%Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}. % also goodman2013knowledge,
The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
\begin{equation}
	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
\end{equation}
where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
%In line with Zipf's Force of Diversification, it is maximized when trees can be decoded fully unambiguously from utterances.

@@

We formalize the cost of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
This is the average of the word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
%The \key{predictability} of the language is the negative entropy of the sentences:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $l$ that belong to the language.
In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. %how many distinct sentences there are.

@@

We first demonstrate that the grammars of real languages are relatively efficient compared to random baselines, providing evidence that grammars evolve towards efficiency.
We compare the efficiency of the actual grammars of the 51 languages from the Universal Dependencies datasets to randomly constructed baseline grammars.
For every one of the 51 languages, we construct ten counterfactual versions by randomly creating ten grammars.
These baseline grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}, see \textit{Materials and Methods} for details).
We estimate predictability and parsability using start-of-the-art machine learning methods (see \textit{Materials and Methods}).
Furthermore, for every one of the 51 languages, we computationally construct grammars that are \emph{optimized} for parsability, predictability, and efficiency (see \textit{Materials and Methods}).

%For every one of the 51 corpora, we computationally construct eight optimal grammars.
%Note that the original orders of the actual languages do not enter this objective.
%See SI for our method for creating optimized grammars.
% For each language, we generated ten baseline word order grammars for the language.
In Figure~\ref{fig:pareto-plane}, we plot predictability and parsability of the grammars of 51 languages, together with the distribution of random baseline grammars, and the Pareto frontier defined by computationally optimized grammars.
Languages are attracted towards the Pareto frontier and away from the region of the baseline languages.
The majority of real languages are above and to the right of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
100~\% of languages improve on either predictability or parsability ($p<0.05$ each, by one-sided $t$-test, with Bonferroni correction).
%By a one-sided \textit{t}-test, 
84~\% of languages improve over the baselines in parsability ($p < 0.05$ each), 78~\% improve in predictability ($p < 0.05$ each).
%At $\lambda = 0.9$, 86~\% of languages improve in overall efficiency ($p<0.05$ each).
%TODO fix bug with czech data

@@

\begin{figure}
    \centering
    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend-viz.pdf}
    \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars (blue distribution). Predictability and parseability scores are $z$-scored within language. The gray curve indicates the Pareto frontier of computationally optimized grammars, averaged over the 51 languages.} % TODO colors, fontsize, etc could conceivably be made prettier
    \label{fig:pareto-plane}
\end{figure}

@@

\begin{figure} % need Efficiency and (maybe) MLE langs on here
%	\includegraphics[width=0.3\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
	\begin{center}
	\includegraphics[width=0.32\textwidth]{../results/correlations/figures/pred-eff-families-2.pdf}
	\end{center}
	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right). Dark blue: Verb patterner \emph{precedes} object patterner. Light blue: Verb patterner \emph{follows} object patterner.}\label{fig:per-lang}
	% efficiency optimization predicts Greenberg's correlations. For every one of the 51 languages, we show which universals are satisfied in the actual grammar (left), and in the grammar optimized for efficiency (right). Efficiency optimization replicates the universals at rates similar to the real grammars.
\end{figure}

>>>>

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics. % \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic

@@

Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study~\cite{lehmann1973structural, vennemann1974theoretical,jackendoff1977x,frazier1985syntactic,chomsky1988language, dryer1992greenbergian, hawkins1994performance}.

A prominent line of research has argued that language universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.

@@

Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between the Forces of \emph{Unification} and \emph{Diversification}.
The Force of Unification is a pressure to reduce the complexity of the language to make production and processing as easy as possible.
Conversely, the Force of Diversification favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
%Minimizing speaker and listener effort represent two opposing forces:
%A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
%A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
These two forces act in opposing directions:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required in order to provide higher informativity.
The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward, lindblom1990explaining, schwartz1997dispersion, hawkins2004efficiency}.

@@

While it has been suggested that these ideas should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult in the domain of grammar, as this requires both large amounts of data representative of language use across languages and cultures, and computational methods for estimating the efficiency of entire languages.
%TODO somewhere start calling this Efficiency
%TODO explain why surprisal. some production/speaker work. (plan reuse)
%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
%This theory blabla Horn RSA
%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
In this work, we address these challenges by combining large-scale text data from 51 languages with state-of-the-art machine learning techiques to estimate the communicative efficiency of grammar.
We apply this approach to the problem of explaining Greenberg's word order correlation universals.
We first show that the word order of languages has evolved to optimize efficiency.
We then show that efficiency optimization accounts for the Greenbergian word order correlations, and makes fine-grained per-language predictions about individual syntactic patterns.
Our results show how general properties of efficient communication gives rise to the universal word order properties of human language.
%We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.

@@

%
%\begin{table}
%	\begin{center}
%%		\setlength\tabcolsep{1.5pt}
%\begin{tabular}{|c|ll|c|cc|ccc}
%	\hline
%	%&
%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \\ 
%	&	verb & object     & & &   \\ 
%	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	Adposition    &    Noun Phrase
%\\
%	&		\emph{to}            & \emph{a friend} &&&\\ \hline
%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
%\\
%	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
% \\
%	&	\emph{has}          & \emph{written}  &&&\\ \hline
%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
%\\
%	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
%\\
%	&	\emph{books} & \emph{that you read}  &&&\\ \hline
%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
%\\
%	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
%\\
%	&	\emph{went} & \emph{to school}  &&&\\ \hline
%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
%\\
%	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
% \hline
%%comp. &
%%\multirow{3}{*}{NP}&
%%	AP &
%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
%\end{tabular}
%	\end{center}
%\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
%	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
%For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
%%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
%%Given the statistical nature of the correlations, not every language satisfies every one of them.
%%Not all correlations are satisfied by every natu
%}\label{table:corr-resu}
%\end{table}
%
%
%
%

\begin{table}
	\begin{center}
%		\setlength\tabcolsep{1.5pt}
\begin{tabular}{|c|ll|ll|}
	\hline
	%&
	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Hindi, ...)} \\ \hline\hline
	& \multicolumn{2}{c}{Correlates with...} & \multicolumn{2}{|c|}{Correlates with...}  \\
	&	Verb & Object     & Object & Verb    \\ \hline\hline
	& kataba & rasaa'il 	&tegami-o & kaita \\
	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    sadiiq       &	tomodachi & ni \\
	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&kaana    &    sadiiq         &	tomodachi & datta \\
	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak & udesho \\
	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& sadiiq  &    John    & 	John no & tomodachi \\
	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uha       & 	anata-ga yonda & hon \\
	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tusil        & 	toochaku suru & koto \\
	&	\emph{that} & \emph{she arrives}  &	\emph{has.arrived} & \emph{that} \\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'ila lmadrasa     & 	gakkoo ni & ittekimashita \\
	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'uriid    &    'an 'ughaadir         & 	yom & itai \\
	& \emph{wants}   &  \emph{to leave}  &	\emph{to leave} & \emph{want} \\ \hline \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples: Across the world, the orders of different constituents are strikingly correlated with that of verb and object. See Materials and Methods.}\label{table:corr-dryer}
\end{table}

@@

%In Information Theory, the usefulness of a communication channel 
The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
%This idea formalizes Zipf's forces of Unification and Diversification
Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.

As we take the underlying messages to be dependency structures, the first term is the degree to which listeners can reconstruct dependency structures from an utterance, i.e., the \key{parseability} if the language. This is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
\begin{equation}
	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
\end{equation}
where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).

We formalize the cost of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
%The \key{predictability} of the language is the negative entropy of the sentences:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $l$ that belong to the language.
In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. %how many distinct sentences there are.

Then, the efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
%The \key{efficiency} of a language is a weighted combination
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
\end{equation}
with an interpolation weight $\lambda \in [0,1)$ (See SI appendix section 5 for mathematical justification). %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).

@@

%In our model, both aspects of efficiency will depend on the grammar:
%Grammars differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
%This reflects the \key{parseability} of the language.
%%Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}. % also goodman2013knowledge,
%The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
%\begin{equation}
%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
%\end{equation}
%where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
%This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
%In line with Zipf's Force of Diversification, it is maximized when trees can be decoded fully unambiguously from utterances.

@@

@@

We first demonstrate that the grammars of real languages are relatively efficient compared to random baselines, providing evidence that grammars evolve towards efficiency.
We compare the efficiency of the actual grammars of the 51 languages from the Universal Dependencies datasets to randomly constructed baseline grammars.
For every one of the 51 languages, we construct 20 counterfactual versions by randomly creating 20 grammars.
These baseline grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}, see \textit{Materials and Methods} for details).
We estimate predictability and parsability using start-of-the-art machine learning methods (see \textit{Materials and Methods}).
Furthermore, for every one of the 51 languages, we computationally construct grammars that are \emph{optimized} for parsability, predictability, and efficiency (see \textit{Materials and Methods}).
For each grammar, we report predictability and parsability as estimated on the data resulting from ordering the dependency structures from the corpus according to the grammar.

%For every one of the 51 corpora, we computationally construct eight optimal grammars.
%Note that the original orders of the actual languages do not enter this objective.
%See SI for our method for creating optimized grammars.
% For each language, we generated ten baseline word order grammars for the language.
In Figure~\ref{fig:pareto-plane}, we plot predictability and parsability of the grammars of 51 languages, together with the distribution of random baseline grammars, and the Pareto frontier defined by computationally optimized grammars.
Languages are attracted towards the Pareto frontier and away from the region of the baseline languages.
The majority of real languages are above and to the right of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
100~\% of languages improve on either predictability or parsability ($p<0.05$ each, by one-sided $t$-test, with Bonferroni correction).
%By a one-sided \textit{t}-test, 
92~\% of languages improve over the baselines in parsability ($p < 0.05$ each), 84~\% improve in predictability ($p < 0.05$ each).
%At $\lambda = 0.9$, 86~\% of languages improve in overall efficiency ($p<0.05$ each).
%TODO fix bug with czech data

@@

\begin{figure}
    \centering
    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend-viz-10.pdf}
    \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars (blue distribution). Predictability and parseability scores are $z$-scored within language. The gray curve indicates the Pareto frontier of computationally optimized grammars, averaged over the 51 languages.} % TODO colors, fontsize, etc could conceivably be made prettier
    \label{fig:pareto-plane}
\end{figure}

@@

\begin{figure} % need Efficiency and (maybe) MLE langs on here
%	\includegraphics[width=0.3\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
	\begin{center}
	\includegraphics[width=0.32\textwidth]{../results/correlations/figures/pred-eff-families-2.pdf}
	\end{center}
	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right). Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...). Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).}\label{fig:per-lang}
	% efficiency optimization predicts Greenberg's correlations. For every one of the 51 languages, we show which universals are satisfied in the actual grammar (left), and in the grammar optimized for efficiency (right). Efficiency optimization replicates the universals at rates similar to the real grammars.
\end{figure}
%\begin{tabular}{|c|ll|c|cc|ccc}
%				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
%\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
%	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
%For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
%%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
%}\label{table:corr-resu}
